---
title: "Chapter 2 - end of chapter questions"
author: "Deborah Hawkshaw"
date: "2023-05-12"
output: html_document
---

# Chapter 2: Small worlds and large worlds

## In-text code
Code is from Chapter 2 in the text book - but has additionally comments etc.
Note: Object names are repeated in different chunks of code for ease.

### Grid approximation

```{r}
# define grid
p_grid <- seq(from = 0, to = 1, length.out = 20) # grid of 20 parameter values

# define prior
prior <- rep(1, 20) # means each parameter value is equally likely

# compute likelihood at each value in grid

likelihood <- dbinom(6, size = 9, prob = p_grid) # 6 Waters, 9 sample size, Each possilble of parameter values

# compute product of likelihood and prior

unstd.posterior <- likelihood * prior

# standardized posterior
posterior <- unstd.posterior / sum(unstd.posterior)

plot(p_grid, posterior, type = "b", xlab = "probability of water")
mtext("20 points")

# Sparser grid ------------------------------

# define grid
p_grid <- seq(from = 0, to = 1, length.out = 5) # grid of 20 parameter values

# define prior
prior <- rep(1, 5) # means each parameter value is equally likely

# compute likelihood at each value in grid

likelihood <- dbinom(6, size = 9, prob = p_grid) # 6 Waters, 9 sample size, Each possilble of parameter values

# compute product of likelihood and prior

unstd.posterior <- likelihood * prior

# standardized posterior
posterior <- unstd.posterior / sum(unstd.posterior)

plot(p_grid, posterior, type = "b", xlab = "probability of water")
mtext("5 points")

# Denser grid ------------------------------

# define grid
p_grid <- seq(from = 0, to = 1, length.out = 100) # grid of 20 parameter values

# define prior
prior <- rep(1, 5) # means each parameter value is equally likely

# compute likelihood at each value in grid

likelihood <- dbinom(6, size = 9, prob = p_grid) # 6 Waters, 9 sample size, Each possilble of parameter values

# compute product of likelihood and prior

unstd.posterior <- likelihood * prior

# standardized posterior
posterior <- unstd.posterior / sum(unstd.posterior)

plot(p_grid, posterior, type = "b", xlab = "probability of water")
mtext("100 points")

# Changing priors ------------------------------

# define grid
p_grid <- seq(from = 0, to = 1, length.out = 20) # grid of 20 parameter values

# define prior
prior <- ifelse(p_grid < 0.5, 0, 1)

# compute likelihood at each value in grid

likelihood <- dbinom(6, size = 9, prob = p_grid) # 6 Waters, 9 sample size, Each possilble of parameter values

# compute product of likelihood and prior

unstd.posterior <- likelihood * prior

# standardized posterior
posterior <- unstd.posterior / sum(unstd.posterior)

plot(p_grid, posterior, type = "b", xlab = "probability of water")
mtext("20 points different parameter values less the 0.5 not possible")

# Changing priors ------------------------------

# define grid
p_grid <- seq(from = 0, to = 1, length.out = 20) # grid of 20 parameter values

# define prior
prior <- exp(-5 * abs(p_grid - 0.5))

# compute likelihood at each value in grid

likelihood <- dbinom(6, size = 9, prob = p_grid) # 6 Waters, 9 sample size, Each possilble of parameter values

# compute product of likelihood and prior

unstd.posterior <- likelihood * prior

# standardized posterior
posterior <- unstd.posterior / sum(unstd.posterior)

plot(p_grid, posterior, type = "b", xlab = "probability of water")
mtext("20 points parmeter values around 0.5 more likely that at tails")

```

### Quadratic approximation

```{r}
library(rethinking)
globe.qa <- quap(
  alist(
    W ~ dbinom(W + L, p), # binomial likelihood (Water follows a binomial distribution of N = W + L with probability p)
    p ~ dunif(0, 1) # uniform prior (all parameter values are equally likely )
  ) 
  ,
  data = list(W = 6, L = 3) # sample data where there was 6 Waters and 3 Lands
)

# display summary of quadratic approximation
precis(globe.qa)

# analytically calculation
W <- 6
L <- 3
curve(dbeta(x, W + 1, L + 1), from = 0, to = 1)

# results from the quadratic approximation
curve(dnorm(x, 0.67, 0.16), lty = 2, add = TRUE)

```
### Markov Chain Monte Carlo

```{r}
n_samples <- 1000
p <- rep(NA, n_samples)
p[1] <- 0.5
W <- 6
L <- 3

for (i in 2:n_samples) {
  p_new <- rnorm(1, p[i - 1], 0.1)
  if (p_new < 0) p_new <- abs(p_new)
  if (p_new > 1) p_new <- (2 - p_new)
  q0 <- dbinom(W, W + L, p[i - 1])
  q1 <- dbinom(W, W + L, p_new)
  p[i] <- ifelse(runif(1) < q1 / q0, p_new, p[i - 1])
}

dens(p, xlim = c(0, 1))
curve(dbeta(x, W + 1, L + 1), lty = 2, add = TRUE)
```

## Practice questions

2E1. (2) Pr(rain|Monday) and (4) Pr(rain, Monday)/Pr(Monday)
(4) because the joint probability, Pr(rain, Monday) = Pr(rain|Monday)*Pr(Monday) so if we solve for Pr(rain|Monday) we get Pr(rain, Monday)/Pr(Monday)

2E2. (3) The probability of Monday given its raining. 

2E3. (1) Pr(Monday|rain) and (4) Pr(rain|Monday)Pr(Monday)/Pr(rain)

2E4. The probability of the earth being covered in 70% water is more likely than all other percentages.

2M1.
```{r}
#(1) W, W, W
p_grid_1 <- seq(0, 1, length.out = 20)
prior_1 <- rep(1, 20)

likelihood_1 <- dbinom(3, size = 3, prob = p_grid_1)
unstd.posterior_1 <- likelihood_1*prior_1
posterior_1 <- unstd.posterior_1/sum(unstd.posterior_1)

plot(p_grid_1, posterior_1, type = "b", xlab = "probability of water")
mtext("W, W, W")

#(2) W, W, W, L
p_grid_2 <- seq(0, 1, length.out = 20)
prior_2 <- rep(1, 20)

likelihood_2 <- dbinom(3, size = 4, prob = p_grid_2)
unstd.posterior_2 <- likelihood_2*prior_2
posterior_2 <- unstd.posterior_2/sum(unstd.posterior_2)

plot(p_grid_2, posterior_2, type = "b", xlab = "probability of water")
mtext("W, W, W, L")

#(3) L, W, W, L, W, W, W
p_grid_3 <- seq(0, 1, length.out = 20)
prior_3 <- rep(1, 20)

likelihood_3 <- dbinom(5, size = 7, prob = p_grid_3)
unstd.posterior_3 <- likelihood_3*prior_3
posterior_3 <- unstd.posterior_3/sum(unstd.posterior_3)

plot(p_grid_3, posterior_3, type = "b", xlab = "probability of water")
mtext("L, W, W, L, W, W, W")
```
2M2.
```{r}
#(1) W, W, W
p_grid_4 <- seq(0, 1, length.out = 20)
prior_4 <- ifelse(p_grid_4 < 0.5, 0, 1)

likelihood_4 <- dbinom(3, size = 3, prob = p_grid_4)
unstd.posterior_4 <- likelihood_4*prior_4
posterior_4 <- unstd.posterior_4/sum(unstd.posterior_4)

plot(p_grid_4, posterior_4, type = "b", xlab = "probability of water")
mtext("W, W, W")

#(2) W, W, W, L
p_grid_5 <- seq(0, 1, length.out = 20)
prior_5 <- ifelse(p_grid_5 < 0.5, 0, 1)

likelihood_5 <- dbinom(3, size = 4, prob = p_grid_5)
unstd.posterior_5 <- likelihood_5*prior_5
posterior_5 <- unstd.posterior_5/sum(unstd.posterior_5)

plot(p_grid_5, posterior_5, type = "b", xlab = "probability of water")
mtext("W, W, W, L")

#(3) L, W, W, L, W, W, W
p_grid_6 <- seq(0, 1, length.out = 20)
prior_6 <- ifelse(p_grid_6 < 0.5, 0, 1)

likelihood_6 <- dbinom(5, size = 7, prob = p_grid_6)
unstd.posterior_6 <- likelihood_6*prior_6
posterior_6 <- unstd.posterior_6/sum(unstd.posterior_6)

plot(p_grid_6, posterior_6, type = "b", xlab = "probability of water")
mtext("L, W, W, L, W, W, W")
```
2M3. How I thought of it suppose earth and mars are made up of 10 marbles each. Earth has 3 white and 7 blue, Mars has 10 white.
If we draw Mars there are 10 ways to get white, if we draw Earth there are 3 ways to get 3. 
Since Mars and Earth are equally likely to pick each marble is equally likely to pick and there are 13 was to get white in a bag full of that composition (20 marbles 13 white, 7 blue). 3/13 is 0.23
OR there is a 30% chance of drawing land if you draw Earth, and there is a 50% chance of drawing Earth, in total the probability of drawing Land is the probability of getting earth and drawing a land probability of drawing mars and getting a land. 
Pr(Earth|land) =  Pr(land|Earth)Pr(Earth)/Pr(Land)
                    =  (0.3*0.5)/(0.3*0.5 + 1*0.5)
                    = 0.23
